---
title: "Females per Roman province per century"
date: "`r format(Sys.time(), '%d %B %Y')`"
author: "by jaro"
output:
  html_document:
    theme: united
    highlight: tango
    code_folding: none
    keep_md: true
  pdf_document: default
---

```{r setup, echo=FALSE, message=FALSE}
knitr::opts_chunk$set(echo=TRUE,error=TRUE)
library(sdam)
load(file="PATH/edhatt.rda")
load(file="PATH/edh_ppd.rda")
```

<div style="margin-bottom:60px;"> </div>


### Package `sdam`
First unistall `sdam` from local machine, and then install package v.0.3.7 from GitHub repo.

```{r sdam, echo=TRUE, eval=FALSE}
# remove sdam
remove.packages("sdam", lib="C:/Users/AUID/R/win-library/version/")
# install from GH
require("devtools")
devtools::install_github("mplex/cedhar", subdir="pkg/sdam")
```

<div style="margin-bottom:30px;"> </div>

Load and check version

```{r}
library(sdam)
packageVersion("sdam")
```



<div style="margin-bottom:60px;"> </div>


### EDH dataset from Vojtech shared folder


```{r edhs, size="tiny", echo=TRUE, eval=FALSE}
# request JSON file from SDDK
edhs = SDDK("EDH_attrs_cleaned_2020-09-30.json", path="/sharingin/648597@au.dk/SDAM_root/SDAM_data/EDH/public", method="GET")
# then SAVE json file on local machine (CHANGE PATHs)
write(edhs, file="PATH/edhs.json")
# transform json file from local machine into list object
edhatt <- get.edhw(file="PATH/edhs.json")
```

### EDH variables
Variables of interest are people's age, gender, along with province labels and dates.

```{r edhvars}
# EDH variables: people (age gender & name) province dates
sel <-c("name", "age: years", "gender")
vars <- c("people", "province_label", "type_of_inscription", "not_after", "not_before")
```

<div style="margin-bottom:30px;"> </div>

Then "select" variables from `edhatt`, which is a list object of the `edhs.json` file. 

```{r, echo=TRUE, eval=FALSE}
# data frame with selected variables in Vojtech shared folder dataset
edh_ppd <- edhw(x=edhatt, select=sel, vars=vars, as="df")
```


<div style="margin-bottom:60px;"> </div>


### Different flavours for selecting variables in data frames

```{r, echo=TRUE, eval=FALSE}
# sdam
sdam::edhw(x=edh_ppd, province="Ach", gender="female")
# dplyr
require("dplyr")
dplyr::filter(edh_ppd, province_label=="Achaia" & gender=="female")
# sqldf
require("sqldf")
sqldf::sqldf('SELECT *
             FROM edh_ppd 
             WHERE province_label="Achaia" AND gender="female"')
```


<div style="margin-bottom:60px;"> </div>


### Aoristic sums Females per Roman province per century

```{r, aors, warning=FALSE}
# Roman provinces
data("rp")
require("dplyr")
edh_femp <- list()
for(i in names(rp)) {
tmp <- prex(edhw(x=dplyr::filter(edh_ppd, province_label==rp[[i]] & gender=="female")), taq="not_before", tpq="not_after", bins=100)
## with 'sdam'
#tmp <- prex(edhw(x=edh_ppd, province=i, gender="female"), taq="not_before", tpq="not_after", bins=100)
if(isTRUE(length(edh_femp)==0L)==TRUE) {
	edh_femp <- tmp
} else {
	ifelse(is.list(edh_femp)==FALSE, edh_femp<-list(edh_femp, tmp), edh_femp<-c(edh_femp, list(tmp)))
}
}
names(edh_femp) <- rp
```

<div style="margin-bottom:30px;"> </div>

Object `edh_femp` is a list of data frames with the probabilities of existence for females per Roman 
province per century in the EDH dataset from Vojtech shared folder. 


```{r, edh_femp, comment="#"}
# look at first entries
head(edh_femp)
```

<div style="margin-bottom:60px;"> </div>


### Bar plots for results

```{r, barplot}
# plot probability of existence in first entry: "Ach" Achaia
barplot(edh_femp[[1]], main=paste(names(edh_femp)[1]," -  females","per century", sep=" "))
```

<div style="margin-bottom:30px;"> </div>
Or else all provinces as pdf file that is located in `PATH`.


```{r, echo=TRUE, eval=FALSE}
pdf("PATH/FILENAME.pdf")
for(i in seq_len(length(edh_femp))) {
# do not plot if NULL
if(is.null(edh_femp[[i]])==FALSE) {
barplot(edh_femp[[i]], main=paste(names(edh_femp)[i]," -  females","per century", sep=" "))
} else { NA }
}
dev.off()
```


<div style="margin-bottom:60px;"> </div>


