---
title: "Variable extraction from EDH with function edhw()"
date: "`r format(Sys.time(), '%d %B %Y')`"
author: "by jaro"
output:
  html_document:
    theme: united
    highlight: tango
    code_folding: none
    keep_md: true
  pdf_document: default
---

```{r setup, echo=FALSE, message=FALSE}
knitr::opts_chunk$set(echo=TRUE,error=TRUE)
library(sdam)
require(rjson)
```

<div style="margin-bottom:60px;"> </div>


### Background (TODO)

`edhw()` is a wrapper function for manipulation of the EDH dataset, 
which was retrieved from the Epigraphic Database Heidelberg open 
data repository through its API. 


TODO.

<div style="margin-bottom:120px;"> </div>


### Choosing variables

```{r edhs, size="scriptsize", echo=-3}
# choose variables of interest in EDH and record it as a data frame
edhs <- edhw(vars=c("type_of_monument", "language", "material", "country", "findspot_ancient",
               "not_after", "not_before"), as="df", limit=10)
edhs
```
We use a regular expression to remove question marks (not optimal for missing data)

```{r edhsreg, echo=-2}
edhs <- as.data.frame(sapply(edhs, function (x)  as.list(gsub('\\?', '', x)) ))
edhs
```

To look at variables like *countries* we need to unlist.

```{r}
unique(unlist(edhs$country))
```


<div style="margin-bottom:120px;"> </div>


### People
Since version 0.2.4 of `sdam` is possible to choose *people* from the `EDH` dataset with function 
`edhw()`. 
In the following example, we extract `"peolple"` from the first 10 records of `EDH` as a data frame. 


```{r edhp, echo=-2}
edhp <- edhw(vars="people", as="df", limit=10)
edhp
```

<div style="margin-bottom:60px;"> </div>


Or in combination of other variables in `EDH`.

```{r edhsp, echo=-2}
edhsp <- edhw(vars=c("people","type_of_monument"), as="df", limit=10)
edhsp
```

<div style="margin-bottom:60px;"> </div>


Since *people* is another list inside `EDH`, each record can be placed in a single row with the 
`bycols` argument. 

```{r edhpc, echo=-2}
edhp <- edhw(vars="people", as="df", limit=10, bycols=TRUE)
edhp
```






<div style="margin-bottom:120px;"> </div>


#### TODO: Petra's questions 

From the SDAM weekly meeting on (ADD DATE) there are questions that can be addressed with `edhw()`: 

* gender
* age
* status
* relationships


<div style="margin-bottom:60px;"> </div>

```{r petra, echo=TRUE, eval=FALSE}
edhs = request("EDH_attrs_cleaned_2020-09-30.json", path="/sharingin/648597@au.dk/SDAM_root/SDAM_data/EDH/public", method="GET")
edhw(x=rjson::fromJSON(edhs), vars=c("type_of_inscription","people"), as="df")
#Error: C stack usage  467226454 is too close to the limit
```

(if you want to save time use `edhw(x=rjson::fromJSON(edhs), vars=c("type_of_inscription","people"), as="df", limit=4)`)

<div style="margin-bottom:60px;"> </div>


#### EDH Open Data Repository

<div style="margin-bottom:60px;"> </div>

##### Gender

Extract records directly from the open data repository through its API.

```{r thrp, echo=TRUE, eval=TRUE}
# example: people inscriptions in Thracia province
thrp <- edhw(x=get.edh(search="inscriptions", province="Thr"), vars="people", as="df")
```

```{r dim, echo=TRUE, eval=TRUE}
# dimensions of the data frame
dim(thrdf)
```

```{r table, echo=TRUE, eval=TRUE}
# contingency table with frequencies
table(thrp$gender)
```

and plot the histogram.

```{r barplot, echo=TRUE, eval=TRUE}
barplot(table(thrp$gender))
```

<div style="margin-bottom:30px;"> </div>


And for the whole `EDH` dataset the command would be

```{r EDH, echo=TRUE, eval=TRUE}
table(edhw(vars="people", as="df")$gender)
```

Etc.











<div style="margin-bottom:60px;"> </div>


