---
title: "Modelling Temporal Uncertainty"
date: "`r format(Sys.time(), '%d %B %Y')`"
author: ""
output:
  html_document:
    theme: united
    highlight: tango
    code_folding: none
    keep_md: true
  pdf_document: default
---

```{r setup, echo=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, error=TRUE)
library(sdam)
```

<div style="margin-bottom:60px;"> </div>


### Intro


<div style="margin-bottom:60px;"> </div>



#### List example pithouse data (Crema &amp; Kobayashi)
```{r}
load(file="c:/R/AUArts/lists/data/res.RData")
```

```{r}
is(res)
```

```{r}
length(res)
```

<div style="margin-bottom:60px;"> </div>


We check again object's structure
```{r}
str(res)
```

Hence, the dataset in object `res` is made of five named components with data frames 
(where as before names in each component are given after the meta-character `$`).


We can access the names with `attr()` function.
```{r}
attr(res, "names")
```


Class of the object's first component corresponding to Nagano prefecture.
```{r}
is(res[[1]])
```

Since we know that the first component is a data frame, we get the variables from the column names.
```{r}
colnames(res[[1]])
```


<div style="margin-bottom:60px;"> </div>



### Plot counts
A graphical representation of the pithouse counts in Nagano prefecture is given as time series


```{r}
plot(ts(res[[1]][,3]), ylab="Pithouse counts")
```




<div style="margin-bottom:120px;"> </div>


<!-- #### Accesing entries in data frames within a list object -->


### Probability of existence
The probability of existence *p*[*t_i*] of an event *e* in a range of time **OMEGA** is the probabilty distribution

<center>*p*[*t_i*] = **delta** *t_i* / **delta** **tau**</center>

where **tau** is the time span of existence, **delta** is for the duration of a given portion of time *t_i* or of **tau**. 


<div style="margin-bottom:60px;"> </div>


To compute *p*[*t_i*], Crema (p. 447) takes the following example:

```
350 ---- 300 ---- 250
     t1       t2

     342 ---- 288
         tau
```
<div style="font-size:9pt;">(years are BC)</div>



<div style="margin-bottom:60px;"> </div>
Hence, **delta** **tau** in this case is 54.


<div style="margin-bottom:60px;"> </div>


#### Temporal resolution

Temporal resolution **phi symbol** refers to the duration of time blocks. 
When **phi symbol** is fixed then we have *aoristic analysis*.

<div style="margin-bottom:60px;"> </div>

To retrieve the minimum probability *P*[**phi symbol**_**alpha**] for the boundaries of the first and last time-blocks, 
we define **phi symbol**_**alpha** =1 where 

<center>**phi**_**alpha** / **delta**_**tau** = 1/54 = *P*[**phi**_**alpha**] = .018 (approx)</center>

<div style="margin-bottom:60px;"> </div>

Hence, the probability of existence for *t_1*

<center>*P*[*t_1*] = *P*[**phi**_**alpha**] &middot; 42</center>
<center><div style="font-size:9pt;">&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;i.e. 342 - 300</div></center>

<div style="margin-bottom:30px;"> </div>

<center>*P*[*t_1*] = .018 &middot; 42 &nbsp; = &nbsp; .78</center>

<div style="margin-bottom:60px;"> </div>

The probability of existence for for *t_2*

<center>*P*[*t_2*] = *P*[**phi**_**alpha**] &middot; 12</center>
<center><div style="font-size:9pt;">&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;i.e. 300 - 288</div></center>

<div style="margin-bottom:30px;"> </div>

<center>*P*[*t_1*] = .018 &middot; 12 &nbsp; = &nbsp; .22</center>

<div style="margin-bottom:60px;"> </div>

Etc.


#### But
*Assumption*: Any equally long portion of time within the time span **tau** has the same probability of existence. 
Plus others...


<div style="margin-bottom:160px;"> </div>


### Aoristic sum
A proxy for evaluating change in the total counts of events across time is the *aoristic sum*, which is the sum of probabilities for each time block.

The aoristic sum is computed accross events for a single portion of time *t_i*.

```
    t1  t2  t3
     ^   ^   ^
a    |   |   |
b    |   |   |
c    |   |   |
     ^   ^   ^
     aoristic
       sums
```

#### But
* uncertainty is embedded in the results
* provides a broad likelihood of process intensity for single time blocks
* unable to provide probabilistic description when two or more time-blocks are directly compared
* more..?






<div style="margin-bottom:160px;"> </div>


### Rate of change
This refers to transition probabilities like increase/stability/decrease.

<center>Rate of chage &nbsp; = &nbsp; (*e* *t_i*+1 - *e* *t_i*) / **phi symbol**</center>


<div style="margin-bottom:160px;"> </div>


### Monte Carlo methods
*Monte Carlo* refers to the method of generating random numbers from a certain distribution, and 
Crema proposes a Monte Carlo solution to overcome the problems of the aoristic sum. 

For archaeological applications, Monte Carlo methods consists on creating artificial "potential" time series 
within the boundary of existence of **tau**, and then assessing the likelihood of the variable's character state over time. 

Each result has a *location* for the mean, and a *scale* for the standard deviation.


<div style="margin-bottom:60px;"> </div>


#### Bayesian analysis

A *Markov chain* is a sequence of numbers where each number is dependent on the previous number in the sequence. 

The term *Markov chain Monte Carlo* (MCMC) refers to algorithms that serves to represent the posterior distribution, which inherently reveals 
the uncertainty of the parameter estimates. Examples are

* Metropolis-Hastings
* Gibbs
* Importance/rejection sampling (related). 


*Proposal* distributions: trapezoidal, normal or gaussian, uniform.




<div style="margin-bottom:60px;"> </div>



