---
title: "R list object: Basics"
date: "`r format(Sys.time(), '%d %B %Y')`"
author: "by jaro"
output:
  html_document:
    theme: united
    highlight: tango
    code_folding: none
    keep_md: true
  pdf_document: default
---

```{r setup, echo=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, error=TRUE)
library(sdam)
```

<div style="margin-bottom:60px;"> </div>


### Lists in R

Two types of list objects in R

* *lists*  as a datatype from `list()`
* *"dotted" lists* as a datatype from `pairlist()` (not commonly used in R computations)

<div style="margin-bottom:60px;"> </div>

More info on these by typing `?list` on the `R` console.

In the manual, we can see that an object `list` is a generalization of a `vector` object.

<div style="margin-bottom:60px;"> </div>

#### Example of dotted list
```{r}
formals(vector)
```

Is the object from a Class?
```{r}
is(formals(vector))
```


<div style="margin-bottom:60px;"> </div>



### Creating Lists

#### empty list with three *components*
```{r}
L0 <- vector("list", length=3)
L0
```

```{r}
is(L0)
```

We can populate this list component by component

```{r}
L0[[1]] <- sample(1:5, 20, replace=TRUE)
L0[[2]] <- rep(letters[1:5], 4)
L0[[3]] <- rpois(20, 1)
L0
```

##### Naming
```{r}
names(L0) <- c("Item1","Item2","Item3")
L0
```

`names` for assigning labels to each component, and these names are given in the object after 
the meta-character `$`. 



<div style="margin-bottom:60px;"> </div>

##### Creating a list with `list()`
```{r}
L1 <- list(x = sample(1:5, 20, replace=TRUE),
           y = rep(letters[1:5], 4), z = rpois(20, 1))
L1
```

Object Class?
```{r}
is(L1)
```

##### Length?
```{r}
length(L1)
```



##### Ways for accessing the first component
```{r}
L1[["x"]]
L1$x
L1[[1]]
```


##### A sublist with only the first component
```{r}
L1[1]
```


##### Renaming
```{r}
names(L1) <- c("Item1","Item2","Item3")
```
```{r}
L1
```

<div style="margin-bottom:60px;"> </div>

##### Indexing
```{r}
L1$Item1[L1$Item1>2]
```


<div style="margin-bottom:60px;"> </div>

#### Joining lists with concatenation
```{r}
L2 <- list(x=c(1,5,6,7), y=c("apple","orange","melon","grapes"))
c(L1,L2)
```

##### Joining two lists by appending
```{r}
append(L1,L2,after=2)
```

##### Ways for adding elements into a list object
```{r}
L1$Item4 <- c("apple","orange","melon","grapes")
L1[[4]] <- c("apple","orange","melon","grapes"); names(L1)[4] <- c("Item4")
L1[["Item4"]] <- c("apple","orange","melon","grapes")
```
```{r}
L1
```


<div style="margin-bottom:60px;"> </div>


