---
title: "R list object: Data frames in lists"
date: "`r format(Sys.time(), '%d %B %Y')`"
author: "by jaro"
output:
  html_document:
    theme: united
    highlight: tango
    code_folding: none
    keep_md: true
  pdf_document: default
---

```{r setup, echo=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, error=TRUE)
library(sdam)
```

<div style="margin-bottom:60px;"> </div>


### Data frames in lists

List objects are very flexible and many people (particularly in humanities) record data in R as data frames "stacked" into a list object.


<div style="margin-bottom:60px;"> </div>

#### Example Pithouse dataset (Crema &amp; Kobayashi)
```{r}
load(file="c:/R/AUArts/lists/data/res.RData")
```

```{r}
is(res)
```

```{r}
length(res)
```

<div style="margin-bottom:60px;"> </div>

#### Check object's structure
```{r}
str(res)
```

Hence, object `res` is made of five named components that are *data frames*. 
As sais before, names in each component follow the meta-character `$`.


We can access the names with `attr()` function.
```{r}
attr(res, "names")
```

<div style="margin-bottom:60px;"> </div>


Class of object's first component
```{r}
is(res[[1]])
```


Since the first component is a data frame, we look at its *dimensions*.
```{r}
dim(res[[1]])
```


That is, 50 obs. with four attributes, which are labeled in the column names.
```{r}
colnames(res[[1]])
```


<div style="margin-bottom:60px;"> </div>


#### Accessing data frames in lists

For accessing the first entry in the first component data frame
```{r}
res[[1]][1,1]
```

Since there are `Levels` then it is treated as `Factor`.


The third column of the data frame in the first component has the pithouse counts.

```{r}
res[[1]][,3]
```


<div style="margin-bottom:60px;"> </div>


#### Plot

We can plot the counts from `"Nagano"`
```{r}
plot(res[[1]][,3], ylab="count")
```

<!-- Or plot it as time-series object -->
```{r eval=FALSE, echo=FALSE}
plot(ts(res[[1]][,3]), ylab="Pithouse counts")
```




<div style="margin-bottom:60px;"> </div>

