---
title: "R list object: From JSON files"
date: "`r format(Sys.time(), '%d %B %Y')`"
author: "by jaro"
output:
  html_document:
    theme: united
    highlight: tango
    code_folding: none
    keep_md: true
  pdf_document: default
---

```{r setup, echo=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, error=TRUE)
library(sdam)
```

<div style="margin-bottom:60px;"> </div>


### Case: EDH

Importing `JSON` files into R typically implies creating a list object.


<div style="margin-bottom:60px;"> </div>

#### Two approaches: `jsonlite` and `rjson`
```{r}
require(jsonlite)
edh <- jsonlite::fromJSON(file="c:/R/AUArts/lists/Petra/sample.json")
```
```{r}
require(rjson)
Edh <- rjson::fromJSON(file="c:/R/AUArts/lists/Petra/sample.json")
```

<div style="margin-bottom:60px;"> </div>

We look at their class object and structure


```{r}
is(edh)
```

```{r eval=FALSE}
str(edh)
```

```
#'data.frame':   6 obs. of  55 variables:
# $ diplomatic_text                 :List of 6
#  ..$ : chr "D M / NONIAE P F OPTATAE / ET C IVLIO ARTEMONI / PARENTIBVS / LIBERTIS LIBERTABVSQVE / POSTERISQVE EORVM / C IV"| __truncated__
#  ..$ : chr "C SEXTIVS PARIS / QVI VIXIT / ANNIS LXX"
#  ..$ : chr "[ ]VMMIO [ ] / [ ]ISENNA[ ] / [ ] XV[ ] / [ ] / ["
#  ..$ : chr "[ ]AVS[ ]LLA / M PORCI NIGRI SER / DOMINAE VENERI / ARAM POSVIT"
#  ..$ : chr "[ ] L SVCCESSVS / [ ] L L IRENAEVS / [ ] C L TYCHES / [ ]VNT / [ ]RENTI F"
#  ..$ : chr "D M S / / MEMMIA AVCTIN[ ] / AN LXX PIA IN SVIS / H S E S T TE[ ] L"
# $ literature                      :List of 6
#  ..$ : chr "AE 1983, 0192.; M. Annecchino, Puteoli 4/5, 1980/81, 286-287, Nr. 17; fig. 17. (C) - AE 1983."
```



<div style="margin-bottom:60px;"> </div>

```{r}
is(Edh)
```

```{r eval=FALSE}
str(Edh)
```

```
#List of 6
# $ :List of 55
#  ..$ diplomatic_text                 : chr "D M / NONIAE P F OPTATAE / ET C IVLIO ARTEMONI / PARENTIBVS / LIBERTIS LIBERTABVSQVE / POSTERISQVE EORVM / C IV"| __truncated__
#  ..$ literature                      : chr "AE 1983, 0192.; M. Annecchino, Puteoli 4/5, 1980/81, 286-287, Nr. 17; fig. 17. (C) - AE 1983."
#  ..$ trismegistos_uri                : chr "https://www.trismegistos.org/text/251193"
#  ..$ id                              : chr "HD000001"
#
```


<div style="margin-bottom:60px;"> </div>

### Selection of relevant attributes

This code is from Petra's `Dealing_with_lists_R_script.R`
```
date_info <- EDH %>% 
  select(id, not_before, not_after, commentary, origdate_text, year_of_find) 
date_info 
```

<div style="margin-bottom:60px;"> </div>


#### Function `edhw()`

With the last version from `edhw()` function from `sdam` (version 0.2.4) you can retrieve that selection from a list object generated either by `rjson` or `jsonlite`.

<div style="margin-bottom:60px;"> </div>

##### Output from `jsonlite`
```{r}
require(sdam)
edhw(x=edh, vars=c('id', 'not_before', 'not_after', 'commentary', 'origdate_text', 'year_of_find'))
```

<div style="margin-bottom:60px;"> </div>

#### Data frame transformation
Argument `as` in `edhw()` allows transforming the list into a data frame where all missing or empty information is treated as *not available* `<NA>`.

```{r}
edhw(x=edh, vars=c('id', 'not_before', 'not_after', 'commentary', 'origdate_text', 'year_of_find'), as="df")
```

<div style="margin-bottom:60px;"> </div>

##### Output from `rjson`

And for `Edh` that is a `vector` `list`, the transformation works as well.

```{r eval=FALSE}
edhw(x=Edh, vars=c('id', 'not_before', 'not_after', 'commentary', 'origdate_text', 'year_of_find'))
edhw(x=Edh, vars=c('id', 'not_before', 'not_after', 'commentary', 'origdate_text', 'year_of_find'), as="df")
```

<div style="margin-bottom:60px;"> </div>


##### TODO

Next steps are:

* Allow selecting all variables/attributes from the list object with `edhw()`
* Incorporate `"People"` in the selection.



<div style="margin-bottom:160px;"> </div>



##### SQL for selection in data frames

Given these two objects:

```{r eval=FALSE}
df1 = data.frame(CustomerId = c(1:6), Product = c(rep("Toaster", 3), rep("Radio", 3)))
df2 = data.frame(CustomerId = c(2, 4, 6), State = c(rep("Alabama", 2), rep("Ohio", 1)))
```

<div style="margin-bottom:60px;"> </div>


*Inner join* and *left join* (substitute *'right'* for right join).

```{r eval=FALSE}
library(sqldf)

df3 <- sqldf("SELECT CustomerId, Product, State 
              FROM df1
              JOIN df2 USING(CustomerID)")

df4 <- sqldf("SELECT CustomerId, Product, State 
              FROM df1
              LEFT JOIN df2 USING(CustomerID)")
```

<div style="font-size:10pt;">https://stackoverflow.com/questions/1299871/</div>


<div style="margin-bottom:60px;"> </div>


##### Even better

```{r eval=FALSE}
?merge
```


<div style="margin-bottom:60px;"> </div>

