---
title: "Knitting with passwords, Antonio's solutions"
output:
  html_document:
    df_print: paged
---


```{r setup, echo=TRUE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, error=TRUE)
devtools::install_github("sdam-au/sdam")

library(sdam)
library(jsonlite)
library(getPass)
library(tibble)
```

# OPTION 1: Loading data with credentials

## 1. Input your credentials to Sciencedata.dk
```{r, echo = TRUE }
mycred <- c("648560@au.dk", "Iamnotgivingyoumypassword")
```

## 2. Make the request 
```{r, echo = TRUE }
resp = request("EDH_sample.json", path="/sharingin/648597@au.dk/SDAM_root/SDAM_data/EDH/", method="GET", cred=mycred)
```

## 3. Make a list from the request and display the first six records (head)
```{r}
edh <- tibble::as_tibble(jsonlite::fromJSON(resp))
head(edh)
```
# Outcome
If you do not want the password to be stored in the document when the report is knitted and can be committed to GitHub, this option is not ideal.


# OPTION 2: Loading data with widget prompt window

## 1. To put credentials when login into Sciencedata.dk
```{r, echo = TRUE }
require("tcltk")
```

## 2. Make the request 
```{r, echo = TRUE }
request("EDH_sample.json", path="/sharingin/648597@au.dk/SDAM_root/SDAM_data/EDH/", method="GET")
```
With this option, a widget prompt window pops-up to enter login details.


## 3. Make a list from the request and display the first six records (head)
```{r}
edh <- tibble::as_tibble(jsonlite::fromJSON(resp))
head(edh)
```

# Outcome
This is an option if you do not want the password to be stored in the document when the report is knitted and can be committed to GitHub.




